---
title: "Recall_Indicator"
author: "Mark McDuffie"
date: "2023-11-18"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}


library(dplyr)
library(rstanarm)
library(ggplot2)
library(sandwich)
library(xts)
library(msm)
library(lubridate)

```


```{r}
recall <- read.csv("recall.csv")
new_df <- data.frame(recall$Report.Received.Date, recall$Manufacturer, as.numeric(recall$Potentially.Affected))
head(new_df)

```


```{r}
#making new data frame renaming variables, removing NA values and 
#changing date column before splitting among companies
new_df2 = data.frame("Date" = new_df$recall.Report.Received.Date, "count" = new_df$as.numeric.recall.Potentially.Affected.,"recall.Manufacturer" = new_df$recall.Manufacturer )
new_df2 = na.omit(new_df2)

new_df2$Date <- as.Date(as.character(new_df2$Date),format="%m/%d/%Y")
new_df2$Date <- format(as.Date(new_df2$Date, "%m/%d/%Y"), "20%y-%m-%d")

new_df2$Date <- as.Date(new_df2$Date)
df3 <- new_df2[order(as.Date(new_df2$Date, format="%m/%d/%Y")),]
```
  
```{r}
ford <- subset(new_df2, recall.Manufacturer == "Ford Motor Company")
mb <- subset(new_df2, recall.Manufacturer == "Mercedes-Benz USA, LLC")
toyota <- subset(new_df2, recall.Manufacturer == "Toyota Motor Engineering & Manufacturing")
honda <- subset(new_df2, recall.Manufacturer == "Honda (American Honda Motor Co.)")
vw <- subset(new_df2, recall.Manufacturer == "Volkswagen Group of America, Inc.")
gm <- subset(new_df2, recall.Manufacturer == "General Motors, LLC")

```


```{r}

ggplot(ford, aes(Date,count)) + geom_point(col = "dark blue")+
  ggtitle("Ford") + labs(y= "Recall occences(Vehicles Affected)", x = "Date")
ggplot(mb, aes(Date,count)) + geom_point(col = "dark blue")+
  ggtitle("Mercedes") + labs(y= "Recall occences(Vehicles Affected)", x = "Date")
ggplot(toyota, aes(Date,count)) + geom_point(col = "dark blue")+
  ggtitle("Toyota") + labs(y= "Recall occences(Vehicles Affected)", x = "Date")
ggplot(honda, aes(Date,count)) + geom_point(col = "dark blue")+
  ggtitle("Honda") + labs(y= "Recall occences(Vehicles Affected)", x = "Date")
ggplot(vw, aes(Date,count)) + geom_point(col = "dark blue")+
  ggtitle("VW") + labs(y= "Recall occences(Vehicles Affected)", x = "Date")
ggplot(gm, aes(Date,count)) + geom_point(col = "dark blue")+
  ggtitle("GM") + labs(y= "Recall occences(Vehicles Affected)", x = "Date")


```

```{r}

ford_ts <- xts(ford$count,ford$Date)
mb_ts <- xts(mb$count,mb$Date)
toyota_ts <- xts(toyota$count,toyota$Date)
honda_ts <- xts(honda$count,honda$Date)
vw_ts <- xts(vw$count,vw$Date)
gm_ts <- xts(gm$count,gm$Date)

```






```{r}
ts_merged = merge(ford_ts, mb_ts, toyota_ts, honda_ts, vw_ts, gm_ts)
head(ts_merged)
```


The Poisson distribution is an appropriate model if the following assumptions are true:

y
 is the number of times an event occurs in a given interval and y
 can take values 0,1,2,...
 (non-negative integers);
The occurrence of one event does not affect the probability that a second event will occur. That is, events occur independently.
The average rate at which events occur is independent of any occurrences. For simplicity, this is usually assumed to be constant, but may in practice vary with time.
Two events cannot occur at exactly the same instant; instead, at each very small sub-interval exactly one event either occurs or does not occur1.

https://rpubs.com/franzbischoff/poisson_regression

```{r}

#number of days between first recall event and last recall event in data
days = 2818
total = sum(ford$count)
l = total/days


```


```{r}
df3$days = day(df3$Date)

df4 <- data.frame('Date' = df3$Date, 'Count' = df3$count, 'Company' = df3$recall.Manufacturer)
df4$day <- mday(df4$Date)
df4$month <- month(df4$Date)
df4$year <- year(df4$Date)
head(df4)
```

```{r}
m1 <- glm(Count ~ Company + day + month + year, family="poisson", data=df4)
summary(m1)


```

```{r}
df4$phat <- predict(m1, type = "response")

library(caret)
process <- preProcess(as.data.frame(df4$phat), method=c("range"))

norm_scale <- predict(process, as.data.frame(df4$phat))
df4$phat_scaled <- norm_scale$`df4$phat`
head(df4)
```

```{r}
## order by program and then by math
p <- df4[with(df4, order(Company, year)), ]

## create the plot
ggplot(p, aes(x = year, y = phat_scaled, colour = Company)) +
  geom_point(aes(y = phat_scaled), alpha=.5, position=position_jitter(h=.2)) +
  geom_line(size = 1) +
  labs(x = "Date", y = "Predicted Recall factors")
```

```{r}
recall_final <- data.frame("Date" = df4$Date, "Vehicles Recalled" = df4$Count, "Manufacturer" = df4$Company, "Recall Factor" = df4$phat_scaled)
head(recall_final)

#write.csv(recall_final, "C:\\Users\\markm\\OneDrive\\Desktop\\6242-Homework\\Project\\recall_indicator.csv", row.names=FALSE)
```
